### 2022-02-08 ms_body/manuscript.Rmd
Add BibTex references to MS
Update AICc model (originally picked wrong rm value)

Sutton, G. F.^1^$\dag$; Martin, G.D. ^1,^^2^

@@

Five candidate MaxEnt models were developed using native-range (Australia) training data for *D. rubiformis*. The five models differed in the feature classes and regularisation multiplier settings used by MaxEnt to build the models. The five models were specified with (1) default MaxEnt settings, and four optimised settings configurations that (2) maximised performance based on AUC~test~, (3) AUC~diff~, (4) OR~10~ and (5) AICc (Table S1). The AUC values for these five models ranged were ranged between 0.82 and 0.97, depending on the MaxEnt settings configuration used. As such, all five candidate models  were considered useful and had high predictive accuracy in distinguishing between *D. rubiformis* recorded presences and background points in its native range. While default MaxEnt settings produced the highest discriminatory power when tested in the training area of Australia (AUC = 0.97), models calibrated with tuned settings configurations led to less overfitting (lower AUC~diff~), better omission rates (best approximated the expectation for OR~10~), and more parsimonious models (AICc) (Fig. 2).

## 3.2. Model transferability

When evaluated on independent testing data in South Africa, the MaxEnt models showed relatively high, but variable, predictive accuracy, depending on the MaxEnt settings used to calibrate the model. For all models, the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores (Fig. 3). The default MaxEnt model found no statistically significant relationship between MaxEnt climatic suitability scores and the probability of *D. rubiformis* establishment in South Africa (Table. 1), and displayed low power to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.64 $\pm$ 0.01) (Fig. 4). In contrast, the four models calibrated using optimised model settings configurations all demonstrated a statistically significant positive relationship between MaxEnt suitability scores and the probability of *D. rubiformis* establishment in South Africa (Table. 1). All four tuned models displayed high power to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.83 - 0.84) than the MaxEnt model calibrated with default settings (AUC: 0.64 $\pm$ 0.01) (Fig. 4). The predicted climatic suitability for *D. rubiformis* across South Africa varied depending on the MaxEnt settings used to calibrate the model (Fig. 5). Default MaxEnt settings yielded higher climatic suitability estimates for *D. rubiformis* than the other MaxEnt models with tuned model settings, notably in regions where *D. rubiformis* is absent (e.g. the north-west coastal region bordering Namibia).

Despite the high performance of the MaxEnt models when transferred in space, MESS maps indicated that the MaxEnt models were extrapolating climatic suitability scores for *D. rubiformis* over the vast majority of South Africa (Fig. S1). Indeed, 54 of the 78 field sites surveyed for *D. rubiformis* in South Africa occur in geographic regions where the MaxEnt models were extrapolating. *Dasineura rubiformis* established at 23 of 24 sites in interpolation (96%), while it established at 44 of 54 sites in extrapolation (82%) (Fig. 6). We found no evidence that a site being in extrapolation/interpolation was a statistically significant predictor of *D. rubiformis* establishment in South Africa (X~2~ = 3.40, d.f = 1, *P* = 0.07). Moreover, classifying sites as being in extrapolation/interpolation yielded lower power to discriminate to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.67 $\pm$ 0.02) than all of the MaxEnt models, barring the model with default settings.

@@

Our results indicated that MaxEnt SDM's were able to provide robust estimates of climatic suitability for *D. rubiformis* in South Africa. AUC values greater than 0.80 are typically considered good fitting models (having high predictive power) [@Fielding1997]. When MaxEnt models, calibrated using native-range records for *D. rubiformis* from Australia, were projected over South Africa, they had AUC scores of 0.83 - 0.84, indicative of high power to discriminate between sites where *D. rubiformis* had established in South Africa or not. Additionally, we showed that the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores. This result demonstrates the usefulness of the MaxEnt SDM and that, at least for our case-study, SDM's can be transferred into novel geographic regions, and provide valuable information to guide policy making and the development of management programmes. For example, biological control programmes such as the programme in which *D. rubiformis* is used to curtail the spread of the highly invasive weed *A. mearnsii* expend many resources (time and money) to mass-rear insect populations that are then released into the field. The positive correlation between MaxEnt suitability scores and *D. rubiformis* establishment means that researchers can develop and implement mass-rearing and experimental release programmes for the insect that prioritise releasing the insect at sites that are highly climatically suitable, and thus, more likely the insect will establish viable populations. This may assist in improving the level of control over the target weed and increasing the efficiency of resources expended by increasing the chances of establishing viable insect populations.

@@

<div id="refs" ></div>

>>>>

Sutton, G. F.^1^$\dag$; Martin, G.D.^1,^^2^

@@

Five candidate MaxEnt models were developed using native-range (Australia) training data for *D. rubiformis*. The five models differed in the feature classes and regularisation multiplier settings used by MaxEnt to build the models. The five models were specified with (1) default MaxEnt settings, and four optimised settings configurations that (2) maximised model performance based on AUC~test~, (3) AUC~diff~, (4) OR~10~ and (5) AICc (Table S1). The AUC values for these five models ranged between 0.89 - 0.92, depending on the MaxEnt settings configuration used. As such, all five candidate models  were considered useful and had high predictive accuracy in distinguishing between *D. rubiformis* recorded presences and background points in its native range. While default MaxEnt settings produced high discriminatory power when tested in the training area of Australia (AUC = 0.89), models calibrated with tuned settings configurations led to less overfitting (lower AUC~diff~), better omission rates (best approximated the expectation for OR~10~), and more parsimonious models (AICc) (Fig. 2).

## 3.2. Model performance

When evaluated on independent testing data in South Africa, the MaxEnt models showed relatively high predictive accuracy, irrespective of the MaxEnt settings used to calibrate the model. For all models, the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores (Fig. 3). All five MaxEnt models demonstrated a statistically significant positive relationship between MaxEnt suitability scores and the probability of *D. rubiformis* establishment in South Africa (Table. 1), and high power to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.79 - 0.91) (Fig. 4). The predicted climatic suitability for *D. rubiformis* across South Africa did vary somewhat depending on the MaxEnt settings used to calibrate the model (Fig. 5). All MaxEnt models predicted relatively high climatic suitability for *D. rubiformis* in the far south-west of South Africa and in the southern Cape region. However, the four MaxEnt models calibrated with tuned model settings differed from the default settings model by predicting climatic suitability for *D. rubiformis* along the entire south and east Coast of South Africa.

Despite the high performance of the MaxEnt models when transferred in space, MESS maps indicated that the MaxEnt models were extrapolating climatic suitability scores for *D. rubiformis* over the vast majority of South Africa (Fig. 6a). Indeed, 54 of the 78 field sites surveyed for *D. rubiformis* in South Africa occur in geographic regions where the MaxEnt models were extrapolating. *Dasineura rubiformis* established at 23 of 24 sites in interpolation (96%), while it established at 44 of 54 sites in extrapolation (82%) (Fig. 6b). We found no evidence that a site being in extrapolation/interpolation was a statistically significant predictor of *D. rubiformis* establishment in South Africa (X~2~ = 3.40, d.f = 1, *P* = 0.07). Moreover, classifying sites as being in extrapolation/interpolation yielded lower power to discriminate between sites where *D. rubiformis* had established or not (AUC: 0.63) than all of the MaxEnt models.

@@

Our results indicated that MaxEnt SDM's were able to provide robust estimates of climatic suitability for *D. rubiformis* in South Africa. AUC values greater than 0.75 are typically considered good fitting models (having high predictive power) [@Fielding1997]. When MaxEnt models, calibrated using native-range records for *D. rubiformis* from Australia, were projected over South Africa, they had AUC scores of 0.83 - 0.84, indicative of high power to discriminate between sites where *D. rubiformis* had established in South Africa or not. Additionally, we showed that the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores. This result demonstrates the usefulness of the MaxEnt SDM and that, at least for our case-study, SDM's can be transferred into novel geographic regions, and provide valuable information to guide policy making and the development of management programmes. For example, biological control programmes such as the programme in which *D. rubiformis* is used to curtail the spread of the highly invasive weed *A. mearnsii* expend many resources (time and money) to mass-rear insect populations that are then released into the field. The positive correlation between MaxEnt suitability scores and *D. rubiformis* establishment means that researchers can develop and implement mass-rearing and experimental release programmes for the insect that prioritise releasing the insect at sites that are highly climatically suitable, and thus, more likely the insect will establish viable populations. This may assist in improving the level of control over the target weed and increasing the efficiency of resources expended by increasing the chances of establishing viable insect populations.

@@

<div id="refs"></div>
